---
author: "Cindy and Kaiting"
date: "2025-06-24"
output: html_document
---

```{r, packages, include=FALSE}
#Load packages
library(tidyverse)
library(ggplot2)
library(foreign)
library(dplyr)
library(tidyr)
library(naniar)
library(psych)
library(lavaan) 
library(semPlot)
library(semptools)
library(corrr)
```

```{r, data, include=FALSE, warning=FALSE}
#Read data
dat_wave1 <- read.spss("UKDA-8658-spss/spss/spss28/covid-19_wave1_survey_cls.sav", to.data.frame = TRUE)
dat_wave2 <- read.spss("UKDA-8658-spss/spss/spss28/covid-19_wave2_survey_cls.sav", to.data.frame = TRUE)
dat_wave3 <- read.spss("UKDA-8658-spss/spss/spss28/covid-19_wave3_survey_cls.sav", to.data.frame = TRUE)
```

# Introduction -- introduction of the database and our research question?

## Database

**Waves:**

Wave 1: May 2020 -- at the height of lockdown restrictions

Wave 2: September/October 2020 -- period between the easing of restrictions

Wave 3: February/March 2021 -- during the third UK lockdown

## Research questions

# Data cleaning process

First we read the data from the COVID-19 database from the COVID-19 Survey in Five National Longitudinal Cohort Studies. Data cleaning was done based on our research question. We decided to only the Next Steps cohort from the database, because it is the only cohort that has social support data and were older than 25 years old to have a higher chance of sustaining stable marriage before the COVID outbreak. Only participants reporting cohabiting with their partner are included to investigate their changes in relationship satisfaction in marriage when having to live with the partner for a longer time during the lockdown.

**Step 1:** include participant based on the inclusion criteria above.

```{r, inclusion}
#only include NS participants with cohabiting partner
tar_wave1 <- dat_wave1 %>%
  filter(str_trim(NSID) != "" & CW1_HHNUMWH_1 == "Yes")

tar_wave2 <- dat_wave2 %>%
  filter(str_trim(NSID) != "" & CW2_HHNUMWH_1 == "Yes")

tar_wave3 <- dat_wave3 %>%
  filter(str_trim(NSID) != "" & CW3_HHNUMWH_1 == "Yes")
```

```{r, cleaning, include=FALSE}
#calculate room-people ratio
tar_wave1$CW1_NUMROOMS <- as.numeric(tar_wave1$CW1_NUMROOMS)
tar_wave1$CW1_HHNUM <- as.numeric(tar_wave1$CW1_HHNUM)
tar_wave1 <- tar_wave1 %>%
  mutate(CW1_PRRATIO = CW1_HHNUM/CW1_NUMROOMS) #denominator cannot be 0 so number of rooms will be the denominator, as there are 0 in CW1_HHNUM

tar_wave2$CW2_NUMROOMS <- as.numeric(tar_wave2$CW2_NUMROOMS)
tar_wave2$CW2_HHNUM <- as.numeric(tar_wave2$CW2_HHNUM)
tar_wave2 <- tar_wave2 %>%
  mutate(CW2_PRRATIO = CW2_HHNUM/CW2_NUMROOMS)

tar_wave3$CW3_NUMROOMS <- as.numeric(tar_wave3$CW3_NUMROOMS)
tar_wave3$CW3_HHNUM <- as.numeric(tar_wave3$CW3_HHNUM)
tar_wave3 <- tar_wave3 %>%
  mutate(CW3_PRRATIO = CW3_HHNUM/CW3_NUMROOMS)

#merge data
all_dat <- bind_rows(tar_wave1, tar_wave2, tar_wave3)
all_dat <- all_dat %>%
  arrange(NSID)

#selected variables
clean_dat <- all_dat %>%
  select(NSID, 
         CW1_FINANCIALMAND, CW1_CURPREG, CW1_PRRATIO, CW1_SOCPROV_1, CW1_SOCPROV_2, CW1_SOCPROV_3, CW1_RELSAT,
         CW2_FINANCIALMAND, CW2_CURPREG, CW2_PRRATIO, CW2_SOCPROV_1, CW2_SOCPROV_2, CW2_SOCPROV_3, CW2_RELSAT, 
         CW3_FINANCIALMAND, CW3_CURPREG, CW3_PRRATIO, CW3_SOCPROV_1, CW3_SOCPROV_2, CW3_SOCPROV_3, CW3_RELSAT)

#changing data into wide format
wide_dat <- clean_dat %>%
  group_by(NSID) %>%
  summarise(across(everything(), ~ first(na.omit(.)), .names = "{.col}"), .groups = "drop")

#Re-code data into numeric format
##record the labels and its corresponding numbers and to compare with the one after re-code, in case there are data missing because of this process
tab_before <- lapply(
  c("CW1_FINANCIALMAND", "CW1_CURPREG", "CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3", "CW1_RELSAT",
    "CW2_FINANCIALMAND", "CW2_CURPREG", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3", "CW2_RELSAT", 
    "CW3_FINANCIALMAND", "CW3_CURPREG", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3", "CW3_RELSAT"),
  \(var) table(wide_dat[[var]], useNA = "ifany")
)
##FINANCIALMAND
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_FINANCIALMAND, CW2_FINANCIALMAND, CW3_FINANCIALMAND),
    ~ recode(.,
             "No information" = -8,
             "Not applicable" = -1,
             "I’m much worse off" = 1,
             "I’m a little worse off" = 2,
             "I’m about the same" = 3,
             "I’m a little better off" = 4,
             "I’m much better off" = 5,
             "I'm much worse off" = 1,
             "I'm a little worse off" = 2,
             "I'm about the same" = 3,
             "I'm a little better off" = 4,
             "I'm much better off" = 5
             )
    ))
##CURPREG
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_CURPREG, CW2_CURPREG, CW3_CURPREG),
    ~ recode(.,
             "No information" = -8,
             "Not applicable" = -1,
             "Yes" = 1,
             "No" = 2
             )
    ))
##SOCPROV_1-2
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_SOCPROV_1, CW1_SOCPROV_2, CW2_SOCPROV_1, CW2_SOCPROV_2, CW3_SOCPROV_1, CW3_SOCPROV_2),
    ~ recode(.,
             "No information" = -8,
             "Not applicable" = -1,
             "Very true" = 1,
             "Partly true" = 2,
             "Not true at all" = 3
             )
    ))
##SOCPROV_3 (Reversed)
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_SOCPROV_3, CW2_SOCPROV_3, CW3_SOCPROV_3),
    ~ recode(.,
    "No information" = -8,
    "Not applicable" = -1,
    "Very true" = 3,
    "Partly true" = 2,
    "Not true at all" = 1
    ),
    .names = "{.col}_R"
  ))
wide_dat <- wide_dat %>%
  select(-c(CW1_SOCPROV_3, CW2_SOCPROV_3, CW3_SOCPROV_3))
##RELSAT
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_RELSAT, CW2_RELSAT, CW3_RELSAT),
    ~ recode(.,
             "No information" = -8,
             "Not applicable" = -1,
             "Very Unhappy" = 1,
             "1: Very unhappy" = 1,
             "1" = 1,
             "2" = 2,
             "3" = 3,
             "4" = 4,
             "5" = 5,
             "6" = 6,
             "7" = 7,
             "Very Happy" = 7,
             "7: Very happy" = 7
             )
    ))
##record another time after re-coding the variables
tab_after <- lapply(
  c("CW1_FINANCIALMAND", "CW1_CURPREG", "CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3_R", "CW1_RELSAT",
    "CW2_FINANCIALMAND", "CW2_CURPREG", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3_R", "CW2_RELSAT", 
    "CW3_FINANCIALMAND", "CW3_CURPREG", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3_R", "CW3_RELSAT"),
  \(var) table(wide_dat[[var]], useNA = "ifany")
)
names(tab_before) <- names(tab_after) <- c("CW1_FINANCIALMAND", "CW1_CURPREG", "CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3", "CW1_RELSAT",
                                           "CW2_FINANCIALMAND", "CW2_CURPREG", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3", "CW2_RELSAT", 
                                           "CW3_FINANCIALMAND", "CW3_CURPREG", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3", "CW3_RELSAT")
tab_before
tab_after
```

**Step 2:** Variables selection after merging dataset from different waves

*Variables:*

[COVID-related stressor:]{.underline} financial stressor, pregnancy status, and personal space.

[Protective factor:]{.underline} social support.

[Outcome variable:]{.underline} relationship satisfaction.

Example code:

```{r, varaibles, eval=FALSE}
#calculate room-people ratio
tar_wave1$CW1_NUMROOMS <- as.numeric(tar_wave1$CW1_NUMROOMS)
tar_wave1$CW1_HHNUM <- as.numeric(tar_wave1$CW1_HHNUM)
tar_wave1 <- tar_wave1 %>%
  mutate(CW1_PRRATIO = CW1_HHNUM/CW1_NUMROOMS) #denominator cannot be 0 so number of rooms will be the denominator, as there are 0 in CW1_HHNUM

#merge data
all_dat <- bind_rows(tar_wave1, tar_wave2, tar_wave3)
all_dat <- all_dat %>%
  arrange(NSID)

#selected variables
clean_dat <- all_dat %>%
  select(NSID, 
         CW1_FINANCIALMAND, CW1_CURPREG, CW1_PRRATIO, CW1_SOCPROV_1, CW1_SOCPROV_2, CW1_SOCPROV_3, CW1_PATIENT, CW1_RELSAT,
         CW2_FINANCIALMAND, CW2_CURPREG, CW2_PRRATIO, CW2_SOCPROV_1, CW2_SOCPROV_2, CW2_SOCPROV_3, CW2_PATIENT, CW2_RELSAT, 
         CW3_FINANCIALMAND, CW3_CURPREG, CW3_PRRATIO, CW3_SOCPROV_1, CW3_SOCPROV_2, CW3_SOCPROV_3, CW3_PATIENT, CW3_RELSAT)
```

**Step 3:** changing dataset into wide format for SEM analysis

Example code:

```{r, wide, eval=FALSE}
#changing data into wide format
wide_dat <- clean_dat %>%
  group_by(NSID) %>%
  summarise(across(everything(), ~ first(na.omit(.)), .names = "{.col}"), .groups = "drop")

```

**Step 4:** re-coding data into numeric scales for statistical analysis

Example code:

```{r, recode, eval=FALSE}
tab_before <- lapply(
  c("CW1_FINANCIALMAND", "CW1_CURPREG", "CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3", "CW1_PATIENT", "CW1_RELSAT",
    "CW2_FINANCIALMAND", "CW2_CURPREG", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3", "CW2_PATIENT", "CW2_RELSAT", 
    "CW3_FINANCIALMAND", "CW3_CURPREG", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3", "CW3_PATIENT", "CW3_RELSAT"),
  \(var) table(wide_dat[[var]], useNA = "ifany")
)

##take FINANCIALMAND as an example
wide_dat <- wide_dat %>%
  mutate(across(
    c(CW1_FINANCIALMAND, CW2_FINANCIALMAND, CW3_FINANCIALMAND),
    ~ recode(.,
             "No information" = -8,
             "Not applicable" = -1,
             "I’m much worse off" = 1,
             "I’m a little worse off" = 2,
             "I’m about the same" = 3,
             "I’m a little better off" = 4,
             "I’m much better off" = 5,
             "I'm much worse off" = 1,
             "I'm a little worse off" = 2,
             "I'm about the same" = 3,
             "I'm a little better off" = 4,
             "I'm much better off" = 5
             )
  ))
##record another time after re-coding the variables
tab_after <- lapply(
  c("CW1_FINANCIALMAND", "CW1_CURPREG", "CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3_R", "CW1_PATIENT", "CW1_RELSAT",
    "CW2_FINANCIALMAND", "CW2_CURPREG", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3_R", "CW2_PATIENT", "CW2_RELSAT", 
    "CW3_FINANCIALMAND", "CW3_CURPREG", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3_R", "CW3_PATIENT", "CW3_RELSAT"),
  \(var) table(wide_dat[[var]], useNA = "ifany")
)

```

# Data exploration

```{r, long_format, include=FALSE}
#reshape into long format for data exploration
long_dat <- wide_dat %>%
  pivot_longer(
    cols = c(CW1_FINANCIALMAND, CW1_CURPREG, CW1_PRRATIO, CW1_SOCPROV_1, CW1_SOCPROV_2, CW1_SOCPROV_3_R, CW1_RELSAT,
             CW2_FINANCIALMAND, CW2_CURPREG, CW2_PRRATIO, CW2_SOCPROV_1, CW2_SOCPROV_2, CW2_SOCPROV_3_R, CW2_RELSAT, 
             CW3_FINANCIALMAND, CW3_CURPREG, CW3_PRRATIO, CW3_SOCPROV_1, CW3_SOCPROV_2, CW3_SOCPROV_3_R, CW3_RELSAT),
    names_to = c("wave", ".value"),
    names_pattern = "CW(\\d+)_(.*)"
  ) %>%
  mutate(wave = as.integer(wave))
```

## Missing data

```{r, NAsummary}
#check missing data
##visualisation of missing data in each variable
gg_miss_var(long_dat, facet = wave)
miss_var_summary(wide_dat)
long_dat %>%
  group_by(wave) %>%
  summarise(across(where(is.numeric), ~sum(is.na(.)), .names = "{.col}"))
```

## Descriptive statistics

```{r, descrptive}
long_dat %>% 
  group_by(wave) %>% 
  summarise(
    financial_mean = mean(FINANCIALMAND, na.rm = TRUE), 
    financial_sd = sd(FINANCIALMAND, na.rm = TRUE),
    preg_mean = mean(CURPREG, na.rm = TRUE), 
    preg_sd = sd(CURPREG, na.rm = TRUE), 
    pr_mean = mean(PRRATIO, na.rm = TRUE),
    pr_sd = sd(PRRATIO, na.rm = TRUE), 
    soc1_mean = mean(SOCPROV_1, na.rm = TRUE), 
    soc1_sd = sd(SOCPROV_1, na.rm = TRUE), 
    soc2_mean = mean(SOCPROV_2, na.rm = TRUE), 
    soc2_sd = sd(SOCPROV_2, na.rm = TRUE),
    soc3_mean = mean(SOCPROV_3_R, na.rm = TRUE),
    soc3_sd = sd(SOCPROV_3_R, na.rm = TRUE),
    rels_mean = mean(RELSAT, na.rm = TRUE), 
    rels_sd = sd(RELSAT, na.rm = TRUE)
    )
```

## Relationship satisfaction across waves

```{r, relsattrand}
library(ggplot2)
ggplot(long_dat, aes(x = factor(RELSAT), fill = factor(wave))) +
  geom_bar(position = "dodge") +
  labs(x = "Relationship satisfaction", y = "Count", fill = "Wave") +
  theme_minimal()
```

```{r, relsatprop}
ggplot(long_dat, aes(x = factor(RELSAT), fill = factor(wave))) +
  geom_bar(position = "fill") +
  labs(x = "Relationship satisfaction", y = "Proportion", fill = "Wave") +
  theme_minimal()
```

## Individual changes to examine the longitudinal trajectories

```{r, ind20, warning=FALSE}
#Individual differences from 20 sample in the dataset
set.seed(12345)
sindex <- sample(unique(long_dat$NSID), 20)

long_dat %>% filter(NSID %in% sindex) %>%
ggplot(., aes(wave, RELSAT, col = NSID)) +
  geom_point() +
  facet_wrap(~ NSID, scales = "fixed") +
  labs(x = "Wave", y = "Relationship satisfaction", col = "NSID") +
  guides(col = guide_legend(nrow = 3)) + theme(legend.position = "bottom") 
```

```{r, indall, warning=FALSE}
ggplot(long_dat, aes(wave, RELSAT, col = NSID)) +
   geom_smooth(aes(group = NSID), method = "lm", se = FALSE, alpha = .05,  linewidth = .5,) +
   geom_smooth(aes(group = 1), method = "lm", se = FALSE, colour = "black", linewidth = 1) +
  labs(x = "Wave", y = "Relationship satisfaction", col = "NSID") +
  guides(col = guide_legend(nrow = 3)) + theme(legend.position = "none")
```

## Changes of each variables overtime

```{r, trends}
#Cannot visualise anything related to the protected factor as it is a latent factor in the model. To simply explore it a bit, take the mean of social support
long_dat <- long_dat %>%
  mutate(
    SOCPROV_total = rowMeans(select(., c(SOCPROV_1, SOCPROV_2, SOCPROV_3_R))
  ))

long_dat %>% 
  group_by(wave) %>% 
  summarise(across(c(FINANCIALMAND, CURPREG, PRRATIO, SOCPROV_total, RELSAT), ~mean(., na.rm = TRUE))) %>%
  pivot_longer(-wave, names_to = "variable", values_to = "mean") %>%
  ggplot(aes(x = wave, y = mean, color = variable, group = variable)) +
  geom_line() + geom_point() +
  labs(x = "Wave", y = "Mean score", title = "Variable trends over time") +
  theme_minimal()
```

## Correlation

```{r, correlation}
#correlation
vars <- long_dat %>%
  select(FINANCIALMAND, CURPREG, PRRATIO, SOCPROV_total, RELSAT)
pairs.panels(vars)
```

## Correlation matrix across wave

```{r, corwave}
wide_dat <- wide_dat %>%
  mutate(
    CW1_SOCPROV_total = rowMeans(select(., c(CW1_SOCPROV_1, CW1_SOCPROV_2, CW1_SOCPROV_3_R))),
    CW2_SOCPROV_total = rowMeans(select(., c(CW2_SOCPROV_1, CW2_SOCPROV_2, CW2_SOCPROV_3_R))),
    CW3_SOCPROV_total = rowMeans(select(., c(CW3_SOCPROV_1, CW3_SOCPROV_2, CW3_SOCPROV_3_R)))
  )

wide_dat %>%
  select(CW1_FINANCIALMAND, CW2_FINANCIALMAND, CW3_FINANCIALMAND,
         CW1_SOCPROV_total, CW2_SOCPROV_total, CW3_SOCPROV_total,
         CW1_RELSAT, CW2_RELSAT, CW3_RELSAT) %>%
  correlate() %>%
  shave() %>%
  rplot()
```

# Fit model

```{r, s1_mdl1}
#Simple CLPM suggesting there might be causal relationship between financial stress and later relationship satisfaction

model1 <- '
  # autoregressive paths
  CW2_FINANCIALMAND ~ a1*CW1_FINANCIALMAND
  CW3_FINANCIALMAND ~ a2*CW2_FINANCIALMAND
  CW2_RELSAT ~ b1*CW1_RELSAT
  CW3_RELSAT ~ b2*CW2_RELSAT

  # cross-lagged paths
  CW2_RELSAT ~ c1*CW1_FINANCIALMAND
  CW3_RELSAT ~ c2*CW2_FINANCIALMAND
  CW2_FINANCIALMAND ~ d1*CW1_RELSAT
  CW3_FINANCIALMAND ~ d2*CW2_RELSAT

  # covariances 
  CW1_FINANCIALMAND ~~ CW1_RELSAT
  CW2_FINANCIALMAND ~~ CW2_RELSAT
  CW3_FINANCIALMAND ~~ CW3_RELSAT
' 

fit1 <- sem(model1, data = wide_dat, missing = "ML", meanstructure = TRUE) 
summary(fit1, standardized = TRUE, fit.measures = TRUE) 
# Fit is okay.. CFI good fit but RMSEA is poor fit. 
# Significant autoregressive paths for both financial stress and relationship satisfaction 
# Non-significant cross-lagged paths. 
# Covariances: Consistently correlated with each other within each wave, but do not influence each other longitudinally. 


#model visualisation
layout_mat <- matrix(NA, nrow=2, ncol=3)
layout_mat[1,1] <- "CW1_FINANCIALMAND"
layout_mat[2,1] <- "CW1_RELSAT"
layout_mat[1,2] <- "CW2_FINANCIALMAND"
layout_mat[2,2] <- "CW2_RELSAT"
layout_mat[1,3] <- "CW3_FINANCIALMAND"
layout_mat[2,3] <- "CW3_RELSAT"

#semlabels <- c("T2 Financial Stress", "T3 Financial Stress", "T2 Relationship Satisfaction", "T3 Relationship Satisfaction", "T1 Financial Stress", "T1 Relationship Satisfaction") ## wanted to relabel the nodes but did not work for now

plt1 <- semPaths(fit1,
         layout = layout_mat,
         whatLabels = "est",
         sizeMan = 9,  
         label.cex = 1.2,
         residuals = FALSE,
         intercepts = FALSE,
         edge.label.cex = 1.2,
         nCharNodes = 0
         )

```

## Models that include social support as a mediator

```{r, c1_mdl1}
#Social support as a mediator between financial stress and relationship satisfaction
model11 <- '
# latent factor for social support: 
  CW1_SUPPORT =~ CW1_SOCPROV_1 + CW1_SOCPROV_2 + CW1_SOCPROV_3_R
  CW2_SUPPORT =~ CW2_SOCPROV_1 + CW2_SOCPROV_2 + CW2_SOCPROV_3_R
  CW3_SUPPORT =~ CW3_SOCPROV_1 + CW3_SOCPROV_2 + CW3_SOCPROV_3_R 
  
  # Autoregressive paths
  CW2_FINANCIALMAND ~ a1*CW1_FINANCIALMAND
  CW3_FINANCIALMAND ~ a2*CW2_FINANCIALMAND
  CW2_SUPPORT ~ s1*CW1_SUPPORT
  CW3_SUPPORT ~ s2*CW2_SUPPORT
  CW2_RELSAT ~ b1*CW1_RELSAT
  CW3_RELSAT ~ b2*CW2_RELSAT

  # Cross-lagged mediation paths 
  CW2_SUPPORT ~ c1*CW1_FINANCIALMAND
  CW3_SUPPORT ~ c2*CW2_FINANCIALMAND
  CW2_RELSAT ~ m1*CW1_SUPPORT
  CW3_RELSAT ~ m2*CW2_SUPPORT

  # Direct effects (done in model1?)
  CW2_RELSAT ~ d1*CW1_FINANCIALMAND
  CW3_RELSAT ~ d2*CW2_FINANCIALMAND 
  
# Indirect effects (should I test this?)
  indirect21 := c1 * m1
  indirect32 := c2 * m2

  # Covariances 
  CW1_FINANCIALMAND ~~ CW1_SUPPORT + CW1_RELSAT
  CW2_FINANCIALMAND ~~ CW2_SUPPORT + CW2_RELSAT
  CW3_FINANCIALMAND ~~ CW3_SUPPORT + CW3_RELSAT
  CW1_SUPPORT ~~ CW1_RELSAT
  CW2_SUPPORT ~~ CW2_RELSAT
  CW3_SUPPORT ~~ CW3_RELSAT
' 
#fit model
fit11 <- sem(model11, data = wide_dat, missing = "ML", meanstructure = TRUE) 

#model indexes
summary(fit11, standardized = TRUE, fit.measures = TRUE) 
```

## Compare if constraint the loadings of the latent variables (social support)

```{r, c1_mdl1_lc}
#Social support as a mediator between financial stress and relationship satisfaction
model11_lc <- '
# latent factor for social support: 
  CW1_SUPPORT =~ l1*CW1_SOCPROV_1 + l2*CW1_SOCPROV_2 + l3*CW1_SOCPROV_3_R
  CW2_SUPPORT =~ l1*CW2_SOCPROV_1 + l2*CW2_SOCPROV_2 + l3*CW2_SOCPROV_3_R
  CW3_SUPPORT =~ l1*CW3_SOCPROV_1 + l2*CW3_SOCPROV_2 + l3*CW3_SOCPROV_3_R 
  
  # Autoregressive paths
  CW2_FINANCIALMAND ~ a1*CW1_FINANCIALMAND
  CW3_FINANCIALMAND ~ a2*CW2_FINANCIALMAND
  CW2_SUPPORT ~ s1*CW1_SUPPORT
  CW3_SUPPORT ~ s2*CW2_SUPPORT
  CW2_RELSAT ~ b1*CW1_RELSAT
  CW3_RELSAT ~ b2*CW2_RELSAT

  # Cross-lagged mediation paths 
  CW2_SUPPORT ~ c1*CW1_FINANCIALMAND
  CW3_SUPPORT ~ c2*CW2_FINANCIALMAND
  CW2_RELSAT ~ m1*CW1_SUPPORT
  CW3_RELSAT ~ m2*CW2_SUPPORT

  # Direct effects
  CW2_RELSAT ~ d1*CW1_FINANCIALMAND
  CW3_RELSAT ~ d2*CW2_FINANCIALMAND 
  
# Indirect effects
  indirect21 := c1 * m1
  indirect32 := c2 * m2

  # Covariances 
  CW1_FINANCIALMAND ~~ CW1_SUPPORT + CW1_RELSAT
  CW2_FINANCIALMAND ~~ CW2_SUPPORT + CW2_RELSAT
  CW3_FINANCIALMAND ~~ CW3_SUPPORT + CW3_RELSAT
  CW1_SUPPORT ~~ CW1_RELSAT
  CW2_SUPPORT ~~ CW2_RELSAT
  CW3_SUPPORT ~~ CW3_RELSAT
' 
#fit model
fit11_lc <- sem(model11_lc, data = wide_dat, missing = "ML", meanstructure = TRUE) 

#model indexes
summary(fit11_lc, standardized = TRUE, fit.measures = TRUE) 

#model comparisons
anova(fit11, fit11_lc) ##not significant, so we can use this constrained model to make sure we are measuring the same latent variable across different time points

#setup layout matrix
layout_matrix <- rbind(
  c(NA, "CW1_FINANCIALMAND", NA, NA, "CW2_FINANCIALMAND", NA, NA, "CW3_FINANCIALMAND", NA),
  c(NA, "CW1_SUPPORT", NA, NA, "CW2_SUPPORT", NA, NA, "CW3_SUPPORT", NA),
  c("CW1_SOCPROV_1", "CW1_SOCPROV_2", "CW1_SOCPROV_3_R", "CW2_SOCPROV_1", "CW2_SOCPROV_2", "CW2_SOCPROV_3_R", "CW3_SOCPROV_1", "CW3_SOCPROV_2", "CW3_SOCPROV_3_R"),
  c(NA, "CW1_RELSAT", NA, NA, "CW2_RELSAT", NA, NA, "CW3_RELSAT", NA)
)

#semlabels <- c("T1 Social Support Q1", "T1 Social Support Q2", "T1 Social Support Q3", "T2 Socual Support Q1", "T2 Socual Support Q2", "T2 Socual Support Q3", "T3 Socual Support Q1", "T3 Socual Support Q2", "T3 Socual Support Q3", "T2 Financial Stress", "T3 Financial Stress", "T2 Relationship Satisfaction", "T3 Relationship Satisfaction", "T1 Financial Stress", "T1 Relationship Satisfaction", "T1 Social Support", "T2 Social Support", "T3 Social Support") ## wanted to relabel the nodes but did not work for now

#model visualisation
plt2 <- semPaths(
  fit11_lc,
  layout = layout_matrix,
  whatLabels = "est",
  edge.label.cex = 0.7,
  sizeMan = 7,
  sizeLat = 7,
  label.cex = 1.2,
  edge.color = "black",
  intercepts = FALSE,
  residuals = FALSE,
  nCharNodes = 0,
  mar = c(8, 4, 8, 4),
  exoCov = FALSE,
  title = FALSE,
  style = "lisrel"
)
```
